---
title: "Analysis of Doctoral Degrees Cross the United States"
author: 
  - Kevin Roe
  - Cher Ning-Li
  - Fangning Zhang
  - Jierui Zhan
  - Jingchuan Xu
thanks: "All data, R code, and other files are available in the following GitHub repository: <https://github.com/Kanghyunroe/us-postdocs>"
date: today
date-format: long
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---


```{r}
#| include: false
#| warning: false
#| message: false

#### Loading packages ####
#install.packages("dplyr")
#install.packages("tidyverse")
#install.packages("knitr")

library(dplyr)
library(tidyverse)
library(knitr)

ipums_data <- read_csv("usa_00002.csv")

```

# How to Obtain the Data {#sec-dataoverview}

The data was obtained from IPUMS USA [@data]. After creating the an IPUMS USA account, you first go to the Get Data button on the home page. Then you want to Select Samples and only select ACS 2022 option and deselect all other default options and you submit your sample selection. Under Select Harmonized Variables and Geographic, select STATEICP. Under Demographic choose SEX, EDUC, EDUCD. You then go under View Cart and create a data extract and make sure to create the data extract as a csv. Once you're able to, download and unzip the data. 

Unzip the data and look for a file called "usa_00002.csv". Load the file in the quarto document and work off the document. 

All analyses of this resesarch were conducted using the statistical programming language R [@citeR].

# Overview of Ratio Estimators Approach {#sec-estimators}

The ratio estimators approach first assumes the number of respondents in California then finds the percentage of respondents with a doctorate. Using this ratio, assuming that all states ahve the same percentage of doctorates, we divide the number of doctorates by the ratio to find the estimated number of respondents for each state. 

# Estimates and Actual Number of Respondents {#sec-estimates}

Please find results under @sec-code

# Explanation for Difference {#sec-explanation}

The reason why there is a difference is because the Ratio Estimators Approach assumes that the percentage of doctorates across all 50 states equals the percentage of doctorates in California. California, a highly educated state, is not a good representation of the entire population in terms of the proportion of doctorates. With many research universities and start-ups the proportion of doctorates in California does not represent the entire population. Therefore, because we use the California proportion for all states, it is not a good representation and there will be discrepancies. 


# Code Output {#sec-code}
```{r}
#| echo: false
#| warning: false
#| message: false
#| results: hide

# Find the total number of respondents of each state.
state_counts <- ipums_data %>%
  group_by(STATEICP) %>%
  summarise(total = n()) %>%
  ungroup() %>%
  mutate(STATEICP = factor(STATEICP, levels = unique(ipums_data$STATEICP))) %>%
  arrange(STATEICP)

# Filter to only include the row for California (state code 71)
california_total_respondents <- state_counts %>%
  filter(STATEICP == 71)

# View the result
california_total_respondents
```


```{r}
#| echo: false
#| warning: false

#### Respondents in Each State with Doctoral Degree ####

doctoral_degree_code <- 116 

doctoral_data <- ipums_data %>%
  filter(EDUCD == doctoral_degree_code)

state_doctoral_counts <- doctoral_data %>%
  group_by(STATEICP) %>%
  summarise(doctoral_count = n()) %>%
  as_tibble()

state_doctoral_counts %>%
  kable(
    col.names = c("State Code", "Number of Doctorates"), # Column names
    caption = "Number of Doctoral Degree Holders by State" # Table title
  )
```

```{r}
#| echo: false
#| warning: false

#### Estimated Total Number of Respondents ####
# Define the total number of respondents in California
california_total_respondents <- 391171

# Extract the number of doctoral degree holders for California (state code 71)
california_doctoral_count <- state_doctoral_counts %>%
  filter(STATEICP == 71) %>%
  pull(doctoral_count)

# Calculate the ratio of doctoral degree holders to the total respondents in California
ratio_california <- california_doctoral_count / california_total_respondents

# Estimate the total number of respondents for each state using the California ratio
state_estimates <- state_doctoral_counts %>%
  mutate(estimated_total_respondents = round(doctoral_count / ratio_california))

# Display the results as a table with custom column names
state_estimates %>%
  kable(
    col.names = c("State Code", "Number of Doctorates", "Estimated Total"), 
    caption = "Estimated Total Number of Respondents by State"
  )
```

```{r}
#| echo: false
#| warning: false

#### Comparison Between Estimate and After ####
# Calculate the actual total number of respondents for each state
actual_respondents <- ipums_data %>%
  group_by(STATEICP) %>%
  summarise(actual_total_respondents = n()) %>%
  as_tibble()

# Merge the estimated and actual total respondents for comparison
comparison <- state_estimates %>%
  left_join(actual_respondents, by = "STATEICP") %>%
  mutate(difference = actual_total_respondents - estimated_total_respondents)

# Display the comparison table
comparison %>%
  kable(col.names = c("State Code","Number of Doctorates", 
                      "Estimated Respondents", "Actual Respondents", 
                      "Difference"),
        caption = "Comparison Between Estimated and Actual Respondents by State"
        )
```
# reference
